---
title: "writeup"
author: "Kelly Chen"
date: "9/25/2021"
output: html_document
---



## Introduction

The world series is a best-of-7 match-up between the champions of the American and National Leagues of Major League Baseball. This document is to analyze and explain the probability calculations related to world series.

## Default Settings

Suppose that the Braves and the Yankees are teams competing in the World Series.
Suppose that in any given game, the probability that the Braves win is P and the probability that the Yankees win is 1-P. 

## Question 1

What is the probability that the Braves win the World Series given that P=0.55?
```{r}
library(ggplot2)
library(dplyr)
```
There are 7 teams in the game, so the total match Braves needed is 7 games, and to win the world series, Braves needs to win at least 4 games.

since pbinom produces the at most 3 games winning probability, we need to subtract that probability to get the result of winning at least 4 games probability.

```{r}
1-pbinom(3,7,0.55)
```

Answer: The probability that the Braves win the World Series given P=0.55 is 0.6082878.This is a relatively high winrate, which means that on average the Braves has a high chance of winning the world series given that they have 0.55 winrate for each match in the world series. 

## Question 2

What is the probability that the Braves win the World Series given that P=x?

Assume p is a random probability between 0-1. Here we take 51 values of P to run the simulation to test the probability of Braves wins the world series.

```{r}
p2<-seq(0,1,0.02)
p_win<-rep(NA,100)

#Simulation
for (i in seq_along(p2)){
  p_win[i]<- 1-pbinom(3,7,p2[i])
}
#remove na values in the p_win
p_win<-na.omit(p_win)

#plot p2 and p_win
plot(p2,p_win,xlab='Probability of the Braves winning a head-to-head matchup',ylab='Pr(Win World Series',type='l',main='Probability of winning the World Series')
```

From the plot we could see the probability that the Braves win the World Series given that P=x, so the p is a random variable we produced from simulation. We could see as x increases, the winning probability of the Braves also increases.

## Question 3

Suppose one could change the World Series to be best-of-9 or some other best-of-X series. What is the shortest series length so that P(Braves win World Series|P=.55)≥0.8


We will set the x value for best of x series, we set the maximum value of x to 100.

```{r}
best_x<-seq(1,999,2)
p_win_1<-rep(NA,100)

#simulation
for (i in seq_along(best_x)){
  best_cal<-ceiling(best_x[i]/2)
  p_win_1[i]<- 1-pbinom(best_cal-1,best_x[i],0.55)
} 

#check p_win_1 value
p_win_1<-na.omit(p_win_1)

#create data frame for the two variables
best_data<-data.frame(number_of_series=best_x,win_probability=p_win_1)

#filter winning probability larger than 0.8
best_data %>%
  filter(win_probability>=0.8) %>%
  filter(win_probability==min(win_probability))

```

Answer: The shortest number of series needed is 71. This means that if the world series is to be best-of-71 games, then the Braves could have a winrate higher than 0.8 given that their winrate for each game is 0.55.

## Question 4

What is the shortest series length so that P(Braves win World Series|P=x)≥0.8? This will be a figure (see below) with PB on the x-axis and series length is the y-axis.

```{r}

#set the x value for best of x series, we set the maximum value of x to 100.

best_x<-seq(3,9999,2)
p_x<-seq(0.51,1,0.01)
series_length<-rep(NA,length(p_x))


#simulation
#Each time when we run the simulation, we keep a record of shortest series length for each Pb.

for (i in seq_along(p_x)){
  for (j in seq_along(best_x)){
  best_cal<-ceiling(best_x[j]/2)
  p_win_1<- 1-(pbinom(best_cal-1,best_x[j],p_x[i]))
  if (p_win_1>=0.8){
    shortest<-best_x[j]
    break}
  }
  series_length[i]<-shortest}

#Remove na values of series_length
series_length<-na.omit(series_length)

#Plot the relationship between Pb and shortest series length.
plot(x=p_x,y=series_length,type='l',xlab='Probability of the Braves winning a head-to-head matchup',ylab='Series Length',main='Short series so that P(Win WS given p)>=0.8')

```

From the plot we could see that the shortest series length needed so that P(Braves win World Series|P=x)≥0.8 is when p is over approximately 0.55, after 0.55 the less the series length neede for the Braves to win.

## Question 5

Calculate P(Pb=0.55|Braves win World Series in 7 games) under the assumption that either Pb=0.55 or Pb=0.45. Explain your solution.

We would utilize Bayes rule and the law of total probability to calculate the problem. We know that **P(A|B)=P(B|A)*P(A)/P(B)**
In this problem, we set P(A) as the probability of Pb=0.55, which is P(Pb=0.55).
We set P(B) as the probability of Braves win world series in 7 games.

Since Pb is either 0.55 or 0.45, so the **P(A)=P(Pb=0.55)=0.5**.
Since there are only two conditions where Pb=0.45 or Pb=0.55, so the total probability of winning the world series should be an average of the two condition.
```{r}
#Calculate the winning probability of Braves when situation is either 0.45 or 0.55
p_win_0.55<-dbinom(4,7,0.55)
p_win_0.45<-dbinom(4,7,0.45)

#The average of winning probability assume either situation happens is 1/2.
p_total_win<-(p_win_0.55+p_win_0.45)/2
p_total_win
```
After calculation, we can know that the **P(B)=total probability of winning the world series=0.5**, this is the average winning probability of the Braves assume the probability of each match winrate situation happens is 0.5.

From question 1, we know that **P(B|A)=P(winning the world series|p=0.55)=p_win_0.55**. P(B) Here is the average of two situations that Braves win 4 lose 3 games probability, which is p_total_win.

```{r}

#Assume the probability of either 0.55 or 0.45 is 1/2.
p_a<-0.5
p_b<-0.5

#According to Bayes Rule, calculate p_final
p_final<-p_win_0.55*p_a/p_total_win
p_final
```
Answer: The probability for the problem is 0.55. This means that if the Braves has a winrate of either 0.55 or 0.45, and the probability of each situation happens is 0.5, then given that the Braves win World Series in 7 games, the winrate of the Braves in each game is 0.55.
