---
title: "05-log-transformation"
author: "Kelly Chen"
date: "10/9/2021"
output: html_document
---
## Part 1

**Gamma Distribution **

X ∼ GAMMA(shape = 3, scale = 1)
```{r}
library(ggplot2)
library(dplyr)
#set up x value
x<-seq(0,12,by=0.5)

#default value
shape<-3
scale<-1
x_mean<-shape*scale
x_median<-qgamma(0.5,shape,scale)

#pdf y value
gamma_x<-dgamma(x,shape,scale)

#Plot gamma distribution
plot(x,gamma_x,type='l',main='gamma distribution PDF')

#calculate mean
dgamma(x_mean,shape,scale)

#Mark mean in plot
abline(v=x_mean,col='blue')
text(4,0.05,'Mean',cex=0.75)

#calculate median

med_x<-dgamma(x_median,shape,scale)
med_x

#Mark median in plot
abline(v=x_median,col='red')
text(2,0.10,'Median',cex=0.75)


```

```{r}
#cdf y value
gamma_cdf<-pgamma(x,shape,scale)

#calculate mean for cdf
pgamma(x_mean,shape,scale)

#plot cdf
plot(x,gamma_cdf,type='l',main='gamma distribution CDF')

#mark mean for cdf
abline(v=x_mean,col='blue')
text(2,0.05,'Mean',cex=0.75)

#calculate median for cdf
pgamma(x_median,shape,scale)

#mark median for cdf
abline(v=x_median,col='red')
text(4,0.05,'Median',cex=0.75)

```

**log normal distribution**

X ∼ LOG NORMAL(μ =  − 1, σ = 1)
```{r}
#set up default value
set.seed(100)
meanlog<--1
sdlog<-1

#log normal distribution y value
draw_log<-rlnorm(n=100,meanlog,sdlog)
log_norm_value<-dlnorm(draw_log,meanlog,sdlog)
log_norm_df<-data.frame(draw_log,log_norm_value)

#x-intercept of mean and median
x_log_mean<-mean(draw_log)
x_log_median<-qlnorm(0.5,meanlog,sdlog)


#arrange dataframe
log_norm_df<-log_norm_df %>%
  arrange(by=draw_log) 


#plot log normal distribution PDF
ggplot(log_norm_df,aes(x=draw_log,y=log_norm_value))+geom_line()+geom_vline(xintercept = x_log_mean,col='blue' )+geom_vline(xintercept = x_log_median,col='red')+geom_text(aes(x=1,y=1),label='mean')+theme_classic()+geom_text(aes(x=0.3,label='median',y=0.5))+ggtitle("Log normal distribution PDF")

```

```{r}
#log normal distribution y value CDF
draw_log<-rlnorm(n=1000,meanlog,sdlog)
log_norm_value<-plnorm(draw_log,meanlog,sdlog)
log_norm_df_cdf<-data.frame(draw_log,log_norm_value)


#arrange dataframe
log_norm_df_cdf<-log_norm_df_cdf %>%
  arrange(by=draw_log) 


#plot log normal distribution PDF
ggplot(log_norm_df_cdf,aes(x=draw_log,y=log_norm_value))+geom_line()+geom_vline(xintercept = x_log_mean,col='blue' )+geom_vline(xintercept = x_log_median,col='red')+geom_text(aes(x=1,y=0.7),label='mean')+theme_classic()+geom_text(aes(x=0.3,label='median',y=0.5))+ggtitle("Log normal distribution PDF")+scale_x_continuous(breaks=seq(0,5,1))
```

**Uniform Distribution**

X ∼ UNIFORM(-1, 13)
```{r}
#set up default value
min<--1
max<-13

#create y value for uniform pdf
uni_pdf<-dunif(x,min,max)

#mean of uniform pdf
x_mean<-(min+max)/2

#plot unifrom PDF
plot(x,uni_pdf,type='s',main='Uniform Distribution PDF')

#Mark mean and median of PDF
abline(v=x_mean,col="red")
text(6,0.08,"Median=Mean")

```

```{r}
#create y value for uniform pdf
uni_cdf<-punif(x,min,max)

#mean of uniform pdf
x_mean<-(min+max)/2

#plot uniform PDF
plot(x,uni_cdf,type='l',main='Uniform Distribution CDF')

#Mark mean and median of PDF
abline(v=x_mean,col="red")
text(6,0.05,"Median=Mean")

```
## Log Transformation

**Gamma Distribution**
```{r}
#generate gamma value
rand_gamma<-rgamma(1000,shape,scale)
log_gamma<-log(rand_gamma)

#find mean of log gamma
gamma_mean<-mean(log_gamma)

#find median of log gamma
gamma_median<-median(log_gamma)

#plot PDF for gamma distribution
hist(log_gamma,main='Log transformation Gamma PDF',breaks = 100)

#add mean and median
abline(v=gamma_mean,col='blue')
text(0.75,15,"mean")
abline(v=gamma_median,col="red")
text(1.2,15,"median")

```

```{r}

#find mean of log gamma
gamma_mean<-mean(log_gamma)

#find median of log gamma
gamma_median<-median(log_gamma)

#plot PDF for gamma distribution
plot(ecdf(log_gamma),main="Gamma distribution CDF")

#add mean and median
abline(v=gamma_mean,col='blue')
text(0.75,0.4,"mean")
abline(v=gamma_median,col="red")
text(1.2,0.5,"median")
```

**Log Normal Distribution**

```{r}
logg_normal<-log(draw_log)

#plot PDF for gamma distribution
hist(logg_normal,main="Log Normal transformation PDF",breaks=100)

#add mean and median
abline(v=mean(logg_normal),col='blue')
text(-0.8,10,"mean")
abline(v=median(logg_normal),col="red")
text(-1.5,10,"median")
```
```{r}
#plot log transformation CDF
plot(ecdf(logg_normal),main="log transformation CDF")

#add mean and median
abline(v=mean(logg_normal),col='blue')
text(-0.7,0.5,"mean")
abline(v=median(logg_normal),col="red")
text(-1.5,0.5,"median")
```
**Uniform Distribution**
```{r}
#create log uniform distribution value
log_unif<-log(runif(1000,0,max))

#create pdf
hist(log_unif,breaks=100,main="Log Uniform transformation PDF")

#Mark Mean and Median
abline(v=mean(log_unif),col='blue')
text(1.1,20,"mean")
abline(v=median(log_unif),col="red")
text(2,20,"median")
```

```{r}
#create CDF
plot(ecdf(log_unif),main="Uniform transformation CDF")

#add mean and median
abline(v=mean(log_unif),col='blue')
text(1,0.5,"mean")
abline(v=median(log_unif),col="red")
text(2.2,0.5,"median")
```

## geometric and arithmetric mean

**Gamma Distribution**
```{r}
#Set up function
geo_mean<-function(x){
  geo<-exp(mean(log(x)))
  return(geo)
}

geo_rec<-rep(NA,1000)
ari_rec<-rep(NA,1000)

for (i in seq(1,1000)){
  gamma_list<-rgamma(100,shape,scale)
  geo_rec[i]<-geo_mean(gamma_list)
  ari_rec[i]<-mean(gamma_list)
}

#plot geometric mean and arithmetic mean
plot(geo_rec,ari_rec,main='Gamma Distribution Geo_mean & Ari_mean')
abline(a=0,b=1,col="red")

#histogram of difference
gamma_diff<-geo_rec-ari_rec
hist(gamma_diff,breaks=100,main="Gamma-Difference of Mean")
```

**Log Normal Distribution**
```{r}

geo_rec_1<-rep(NA,1000)
ari_rec_1<-rep(NA,1000)

for (i in seq(1,1000)){
  lognorm_list<-rlnorm(100,meanlog,sdlog)
  geo_rec_1[i]<-geo_mean(lognorm_list)
  ari_rec_1[i]<-mean(lognorm_list)
}

#plot geometric mean and arithmetic mean
plot(geo_rec_1,ari_rec_1,main='Log Normal Distribution Geo_mean & Ari_mean')
abline(a=0,b=1,col="red")

#histogram of difference
lognorm_diff<-geo_rec_1-ari_rec_1
hist(lognorm_diff,breaks=100,main="Log Normal-Difference of Mean")
```

**Uniform Distribution**
```{r}
geo_rec_2<-rep(NA,1000)
ari_rec_2<-rep(NA,1000)

for (i in seq(1,1000)){
  uni_list<-runif(100,min,max)
  geo_rec_2[i]<-geo_mean(uni_list)
  ari_rec_2[i]<-mean(uni_list)
}

#plot geometric mean and arithmetic mean
plot(geo_rec_2,ari_rec_2,main='Uniform Distribution Geo_mean & Ari_mean')
abline(a=0,b=1,col="red")

#histogram of difference
unif_diff<-geo_rec_2-ari_rec_2
hist(unif_diff,breaks=100,main="Uniform-Difference of Mean")
```
## Part 3 

```{r}
rec_1<-rep(NA,1000)
rec_2<-rep(NA,1000)



for (i in seq(1,1000)){
  uni_list<-runif(100,0,10)
  rec_1[i]<-mean(log(uni_list))
  rec_2[i]<-log(mean(uni_list))
}

#plot geometric mean and arithmetic mean
plot(rec_1,rec_2,main='E(Log(X)) VS Log(E(X))')
abline(a=0,b=1,col="red")

#histogram of difference
rec_diff<-rec_1-rec_2
hist(rec_diff,breaks=100,main="Histogram E(Log(X)) VS Log(E(X))")
```

**Summary**
Since rec2 is log (E[X]), and the difference of rec_1-rec_2 is always negative to zero, so rec_2 >= rec_1, which means Log(E(x)) >=E(Log(x)). Log(E(x)) is always larger or equal to the other.
